---
title: "Analysis"
author: "Sneha Verma"
date: "5/1/2022"
output: word_document
purpose: consists of all the steps involved in generating the regression analysis results.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import

```{r}
setwd('/Volumes/GoogleDrive-116166411084487006202/Shared drives/College Graduation Rates/Verma-493-Project')
```


```{r}

library(readxl)

race = read_excel('Analysis Data/race.xlsx')

# Use this code if the line above does not work

# race = read_excel('/Volumes/GoogleDrive-116166411084487006202/Shared drives/College Graduation Rates/Replication Documentation/Analysis Data/race.xlsx')

head(race)
```

```{r}
sex = ('Analysis Data/sex.xlsx')

# Use this code if the line of code above does not work.
# sex = read_excel('/Volumes/GoogleDrive-116166411084487006202/Shared drives/College Graduation Rates/Replication Documentation/Analysis Data/sex.xlsx')

head(sex)
```

```{r}
pell = read_excel('Analysis Data/pell.xlsx')

# Use this line of code if the line above does not work.
# pell = read_excel('/Volumes/GoogleDrive-116166411084487006202/Shared drives/College Graduation Rates/Replication Documentation/Importable Data/Final/pell.xlsx')

head(pell)
```

## Econometric Analysis

### Race

The regression equation (without year) is:

graduation rate = $\beta$$_{0}$ + $\beta$$_{1}$ * Black/African American + $\beta$$_{2}$ * Hispanic/Latino + $\beta$$_{3}$ * Native Hawaiian/Pacific Islander + $\beta$$_{4}$ * Non-resident Alien + $\beta$$_{5}$ * Two or more races + $\beta$$_{6}$ * White + $\beta$$_{8}$ * carleton college + $\beta$$_{9}$ * grinnell college + $\beta$$_{10}$ * lawrence university + $\beta$$_{11}$ * macalester college

```{r}
library(tidyverse)

# First model without year

race_model = lm(grad_rate ~ race + college, data = race)

race_model
```

According to the results, the coefficients are:

graduation rate = 0.86 + (-0.094) * Black/African American + (-0.02) * Hispanic/Latino + (-0.22) * Native Hawaiian/Pacific Islander + (0.024) * Non-resident Alien + (-0.004) * Two or more races + (0.053) * White +  (-0.061) * grinnell college + (-0.29) * lawrence university + (-0.013) * macalester college

```{r}
summary(race_model)
```

The results from the above command was used to format Table 1 in the Final Paper.

The regression equation with year is:

graduation rate = $\beta$$_{0}$ + $\beta$$_{1}$ * Black/African American + $\beta$$_{2}$ * Hispanic/Latino + $\beta$$_{3}$ * Native Hawaiian/Pacific Islander + $\beta$$_{4}$ * Non-resident Alien + $\beta$$_{5}$ * Two or more races + $\beta$$_{6}$ * White +  $\beta$$_{7}$ * year +$\beta$$_{8}$ * carleton college + $\beta$$_{9}$ * grinnell college + $\beta$$_{10}$ * lawrence university + $\beta$$_{11}$ * macalester college

```{r}
# Second model with year

race_model_2 = lm(grad_rate ~ race + year + college, data = race)

race_model_2
```


```{r}
summary(race_model_2)
```

According to the results, the coefficients are:

graduation rate = -6.55 + (-0.094) * Black/African American + (-0.02) * Hispanic/Latino + (-0.23) * Native Hawaiian/Pacific Islander + (0.025) * Non-resident Alien + (-0.005) * Two or more races + (0.054) * White + (0.004) * year + (-0.076) * grinnell college + (-0.30) * lawrence university + (-0.031) * macalester college

Both the models exclude the Asian students and Carleton College.

### Sex

The regression equation without year is:

graduation rate = $\beta$$_{0}$ + $\beta$$_{1}$ * male + $\beta$$_{2}$ * grinnell college + $\beta$$_{3}$ * lawrence university

```{r}
# First model without year

sex_model = lm(grad_rate ~ sex + college, data = sex)

sex_model
```

```{r}
summary(sex_model)
```

The results from the above command was used to format Table 2 in the Final Paper.

According to the results, the coefficients are:

graduation rate = 0.90 + (-0.018) * male + (-0.082) * grinnell college + (-0.27) * lawrence university

The regression equation with year is:

graduation rate = $\beta$$_{0}$ + $\beta$$_{1}$ * male + $\beta$$_{2}$ * year + $\beta$$_{3}$ * grinnell college + $\beta$$_{4}$ * lawrence university

```{r}
# Second model with year

sex_model_2 = lm(grad_rate ~ sex + year + college, data = sex)

sex_model_2
```

```{r}
summary(sex_model_2)
```


According to the results, the coefficients are:

graduation rate = 0.58 + (-0.018) * male + (0.0002) * year + (-0.083) * grinnell college + (-0.27) * lawrence university

These models exclude female students and Carleton College.

### Pell Grant

The regression equation without year is:

graduation rate = $\beta$$_{0}$ + $\beta$$_{1}$ * Received neither + $\beta$$_{2}$ * Received loan only + $\beta$$_{3}$ * coe college + $\beta$$_{4}$ * grinnell college + $\beta$$_{5}$ * lawrence university

```{r}
# First model without year

pell_model = lm(grad_rate ~ pell_status + college, data = pell)

pell_model
```

```{r}
summary(pell_model)
```

The results from the above command was used to format Table 3 in the Final Paper.

According to the results, the coefficients are:

graduation rate = 0.83 + (0.084) * Received neither + (0.063) * Received loan only + (-0.28) * coe college + (-0.072) * grinnell college + (-0.26) * lawrence university

The regression equation with year is:

graduation rate = $\beta$$_{0}$ + $\beta$$_{1}$ * Received neither + $\beta$$_{2}$ * Received loan only + $\beta$$_{3}$ * coe college + $\beta$$_{4}$ * grinnell college + $\beta$$_{5}$ * lawrence university + $\beta$$_{6}$ * year

```{r}
# Second model with year

pell_model_2 = lm(grad_rate ~ pell_status + college + year, data = pell)

pell_model_2
```
```{r}
summary(pell_model_2)
```


According to the results, the coefficients are:

graduation rate = -0.016 + (0.084) * Received neither + (0.063) * Received loan only + (-0.28) * coe college + (-0.074) * grinnell college + (-0.26) * lawrence university + (0.0004) * year

These models exclude received federal pell grant students and Carleton College.
